---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
pdat<-read.csv("activity.csv")

```



## What is mean total number of steps taken per day?
```{r}
sumstep<-tapply(pdat$steps, factor(pdat$date), function(x) sum(x, na.rm=T))
hist(sumstep,5, xlab="number of steps per day", main="no fill for missing data")
meanstep<-mean(sumstep)
medstep<-median(sumstep)
```
The mean number of steps per day is `r meanstep` and the median is `r medstep`.
## What is the average daily activity pattern?
```{r}
mnin<-tapply(pdat$steps, factor(pdat$interval), function (x) mean(x, na.rm=T))
tmint<-levels(factor(pdat$interval))
tmintpad<-sprintf("%04d", as.numeric(tmint))
ptime<-strptime(tmintpad, "%H%M")
plot(ptime, mnin, type="l", xlab="time (5 min intervals)", ylab="average number of steps", main="average daily activity pattern")
maxsteps=ptime[which.max(mnin)]
printsteps<-strftime(maxsteps,"%H:%M")
```
On average, the maximum number of steps per day are taken at `r printsteps` 

## Imputing missing values
```{r}
rowswna<-sum(!complete.cases(pdat)) 
missvect<-(which(!complete.cases(pdat)))
pdatmod<-pdat

tind <- rep(NA, rowswna)
for(i in 1:rowswna){
tind[i]<-which(tmint==pdat$interval[missvect[i]])
}
pdatmod$steps[missvect]<-mnin[tind]

sumstepmod<-tapply(pdatmod$steps, factor(pdatmod$date), function(x) sum(x, na.rm=T))
hist(sumstepmod,5, xlab="number of steps per day", main="missing data filled with average number of steps for that time interval")
meanstepmod<-mean(sumstepmod)
medstepmod<-median(sumstepmod)
options(scipen=5)

```
The number of rows with missing data is `r rowswna`. When we replace the missing data with the overall mean for that particular time interval, the mean number of steps per day is `r meanstepmod` and the median is `r medstepmod`. These values are larger those calculated with missing data. 
## Are there differences in activity patterns between weekdays and weekends?
```{r, results='hide'}
n<-("weekday")
m<-("weekend")
pdatmod$day<-as.factor(weekdays(strptime(pdatmod$date, "%Y-%m-%d" )))
dayvect<-ifelse(pdatmod$day %in% c("Saturday", "Sunday"), m, n)
dayvect<-as.factor(dayvect)
pdatmod$day<-dayvect

```

```{r}
library(lattice) 
wkact<-with(pdatmod, aggregate(steps, list(interval,day), mean))

xyplot(wkact[,3] ~ wkact[,1]|factor(wkact[,2]),layout=c(1,2),
   type="l", xlab="interval", ylab="Average number of steps")

